The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2012-11-19 script/en/chapters/moderncryptography.tex
CT script: Minor changes (more consistent, more up-to-date)

\newpage
%********************************************************************
\hypertarget{Chapter_ModernCryptography}{}
\chapter[The Mathematical Ideas behind Modern Cryptography]
        {The Mathematical Ideas behind Modern Cryptography\footnotemark}
\footnotetext{%
    \index{NT, Learning Tool for Number Theory}%
    \index{Educational tool NT}%
    With the educational tool for number theory {\bf NT} you can apply
    some of the methods introduced here (RSA, Rabin, DH, ElGamal)
    (see learning unit 4.2 and 4.3, pages 9-17/17).\\
    NT can be called in CrypTool\index{CrypTool} via the menu path
    {\bf Indiv. Procedures \textbackslash{} Number Theory Interactive
    \textbackslash{} Learning Tool for Number Theory}.
    See appendix \ref{s:appendix-Learn-NT}.
}
\label{Chapter_ModernCryptography}
%********************************************************************
(Oyono R. / Esslinger B./ Schneider J., Sep. 2000; 
            Updates Nov. 2000, Feb. 2003, Apr. 2007, Mar. 2010)

%\vskip +30 pt
\begin{ctsquote}
   I don't know if its getting better, if we change it,\\
   but I know, that we have to change it, if it should become better.%TODO - Rest von German
\caption[Georg Christoph Lichtenberg]{Georg Christoph Lichtenberg\footnotemark}
\end{ctsquote}
\footnotetext{%
  Georg Christoph Lichtenberg\index{Lichtenberg, Georg Christoph}, 
  German writer and physicist (1742-1799),\\ 
  (also see: \href{http://en.wikipedia.org/wiki/Georg\_Christoph\_Lichtenberg}
              {\tt http://en.wikipedia.org/wiki/Georg\_Christoph\_Lichtenberg})
}

@@

A 0-1 knapsack problem can be solved by testing the $ 2^n $   possibilities for 
$ x_1, \dots, x_n $. The best method requires $ O(2^{n/2}) $  operations, which 
for $ n=100 $  with $ 2^{100} \approx 1.27 \cdot 10^{30} $  and  $ 2^{n/2} 
\approx 1.13 \cdot 10^{15} $ represents an insurmountable hurdle for computers.
However, for special $ a_1, \dots, a_n $   the solution is quite easy to find, 
e.g. for $ a_i = 2^{i-1}. $  The binary representation of $ g $ immediately 
delivers $ x_1, \dots, x_n$. In general, the a 0-1 knapsack problem can be 
solved easily if a \index{Permutation} permutation\footnote{A permutation\index{Permutation} $\pi$ 
of the numbers $1, \dots, n$ is a change in the order in which these numbers are 
listed. For example, a permutation $\pi$ of $(1,2,3)$ is $(3,1,2),$ i.e. 
$\pi(1) = 3$, $\pi(2) =1$ 
and $\pi(3) = 2$.} 
$ \pi $  of $ 1, \dots, n $  exists with $ a_{\pi (j)} > \sum_{i=1}^{j-1} 
a_{\pi(i)} $. If, in addition, $ \pi $ is the identity, i.e. $ \pi(i)=i $ for $ 
i=1,2,\dots,n, $ then the sequence $ a_1, \dots , a_n $ is said to be 
super-increasing.
Crypto procedure~\ref{knapsackalgo} solves the knapsack problem with a super-increasing 
sequence in the time of $ O(n). $
\begin{cryptoprocedure}
\begin{tabbing}
\hspace*{0.5cm} \= \hspace*{0.5cm} \= \hspace*{0.5cm} \= \kill
\>{\bf for} $ i=n $ {\bf to} 1 {\bf do} \\
\>\> {\bf if} $ T\geq a_i $ {\bf then}\\
\>\> \> $ T:=T-s_i $ \\
\>\>\> $ x_i:=1 $ \\
\>\> {\bf else} \\
\>\>\> $ x_i:=0 $ \\
\>{\bf if} $ T=0 $ {\bf then} \\
\>\> $ X:=(x_1, \dots, x_n) $ is the solution. \\
\>{\bf else} \\
\>\> No solution exists.
\end{tabbing}
\caption{Solving knapsack problems with super-increasing weights}
\label{knapsackalgo}
\end{cryptoprocedure}

@@

%--------------------------------------------------------------------
\hypertarget{RSAVerfahren}{}
\subsection[The RSA procedure]
{The RSA procedure\footnote{%
Please compare chapters \ref{rsabeweis}, ff.
}$^,$\footnote{%
Using CrypTool\index{CrypTool} you can gain practical experience
with the RSA procedure via the menu {\bf Indiv.Procedures \textbackslash{} 
RSA Cryptosystem \textbackslash{} RSA Demonstration}.
}}

% LateX-Learnings: Mit \footnotemark statt \footnote
% stand der höhere Index unten zweimal (oben ok).
%
%  {The RSA procedure\footnotemark$^,$\footnotemark}
%  \footnotetext{
%  Please compare chapters \ref{rsabeweis}, ff.
%  }
%  \footnotetext{
%  Using CrypTool\index{CrypTool} you can gain practical experience
%  with the RSA procedure via the menu {\bf Indiv.Procedures \textbackslash{} 
%  RSA Cryptosystem \textbackslash{} RSA Demonstration}.
%  }
%  
%   in the article ``Introduction to Elementary Number Theory with Examples'',
%   chapter~\ref{patternsandstructures}
%   \hyperlink{RSABeweis}{Beweis des RSA-Verfahrens mit Euler-Fermat}

\index{RSA} \label{rsaverfahren}

@@

For the Euler phi function it holds that: 
\[
\phi (1)=1, \phi(2)=1,
\phi(3)=2, \phi (4)=2, \phi(6)=2, \phi (10)= 4, \phi (15)=8.
\]
For example, $ \phi (24)=8, $ because
\[
|\{ x <24 : \ggt (x,24) =1 \}| =|\{1,5,7,11,13,17,19,23\}|.
\]
If  $ p $ is a prime number, then $ \phi (p)= p-1. $ \\
If we know the various prime factors $ p_1, \dots , p_k $ of $ N, $ then 
\[
\phi(N) = N \cdot (1-\frac{1}{p_1}) \,
\cdots \, (1-\frac{1}{p_k}).\footnote{Further formulas for the Euler phi
function are in the article ``Introduction to Elementary Number Theory with
Examples'', chapter~\ref{patternsandstructures}.} %or \ref{L-Euler-Function}
\]
Table~\ref{phi15} gives the values up to $15$. 
In the case of $ N=pq $, 
\[
\phi (N)= pq(1-1/p)(1-1/q) = p(1-1/p)q(1-1/q)=(p-
1)(q-1).
\]
\end{remark}

@@

\vskip +20 pt
%--------------------------------------------------------------------
\hypertarget{HT-Discrete-Logarithm-as-Basis}{}
\section[The discrete logarithm as basis for public key procedures]
           {The discrete logarithm as basis for public key procedures\footnotemark}
\footnotetext{%
    \index{NT, Learning Tool for Number Theory}%
    \index{Educational tool NT}%
    With the educational tool for number theory {\bf NT} you can play with the
    distribution of the discrete logarithm values and apply Shank's
    baby-step-giant-step method:
    See learning units 6.1-6.3, pages 1-6/6.\\
    NT can be called in CrypTool\index{CrypTool} via the menu path
    {\bf Indiv. Procedures \textbackslash{} Number Theory Interactive
    \textbackslash{} Learning Tool for Number Theory}.
    See appendix \ref{s:appendix-Learn-NT}.
}
\label{L-Discrete-Logarithm-as-Basis}
\index{Discrete logarithm}\index{Logarithm problem!discrete}
Discrete logarithms form the basis for a large
number of algorithms for public-key procedures.

@@

The current record\index{Logarithm problem!record} (as of April 2007) for calculating discrete logarithms was
established in February 2007 by the group Kleinjung, Franke and Bahr at
University of Bonn.\footnote{%
   \href{http://www.nabble.com/Discrete-logarithms-in-GF(p)-----160-digits-t3175622.html}
{\texttt{http://www.nabble.com/Discrete-logarithms-in-GF(p)-----160-digits-t3175622.html}}}
Kleinjung calculated the discrete logarithm modulo a 160 digit prime number $p$
and generator $g$:
$$
\begin{array}{r@{\:}c@{\:}l}
p & = & \lfloor 10^{159}\pi\rfloor + 119849 \\
  & = & 314159265358979323846264338327950288419716939937510582097494 \\
  &   & 459230781640628620899862803482534211706798214808651328230664 \\
  &   & 7093844609550582231725359408128481237299\\
g & = & 2
\end{array}
$$
The discrete logarithms $k$ of the following integer $y$ was
determined:\footnote{The integer $y$ was chosen as the first 159 digits of the
Euler number $e$.}
$$
\begin{array}{r@{\:}c@{\:}l}
y & = & \lfloor 10^{159}e\rfloor \\
  & = & 271828182845904523536028747135266249775724709369995957496696 \\
  &   & 762772407663035354759457138217852516642742746639193200305992 \\
  &   & 1817413596629043572900334295260595630738 \\
k & = & \log_g(y)\mod p\\
  & = & 829897164650348970518646802640757844024961469323126472198531 \\
&   & 845186895984026448342666252850466126881437617381653942624307 \\
&   & 537679319636711561053526082423513665596
\end{array}
$$
The search was performed with GNFS method (General Number Field Sieve,
Index-Calculus) \index{General Number Field Sieve (GNFS)} and took about 17 CPU
years on 3.2 GHz Xeon machines.

%--------------------------------------------------------------------
\subsection[Diffie-Hellman key agreement]
{Diffie-Hellman key agreement\footnotemark}
\footnotetext{%
With CrypTool\index{CrypTool} this exchange protocol has been
visualized: you can execute the single steps with concrete numbers using 
menu {\bf Indiv. Procedures \textbackslash{} Protocols \textbackslash{} Diffie-Hellman Demonstration}.
}

>>>>

\newpage
%********************************************************************
\hypertarget{Chapter_ModernCryptography}{}
\chapter[The Mathematical Ideas behind Modern Cryptography]
        {The Mathematical Ideas behind Modern Cryptography\footnotemark}
\footnotetext{%
    \index{NT, Learning Tool for Number Theory}%
    \index{Educational tool NT}%
    With the educational tool for number theory {\bf NT} you can apply
    some of the methods introduced here (RSA, Rabin, DH, ElGamal)
    (see learning unit 4.2 and 4.3, pages 9-17/17).\\
    NT can be called in CT1\index{CrypTool 1} via the menu path
    {\bf Indiv. Procedures \textbackslash{} Number Theory Interactive
    \textbackslash{} Learning Tool for Number Theory}.
    See appendix \ref{s:appendix-Learn-NT}.
}
\label{Chapter_ModernCryptography}
%********************************************************************
(Oyono R. / Esslinger B./ Schneider J., Sep. 2000; 
            Updates Nov. 2000, Feb. 2003, Apr. 2007, Mar. 2010)

%\vskip +30 pt
\begin{ctsquote}
   I don't know if its getting better, if we change it,\\
   but I know, that we have to change it, if it should become better.%TODO - Rest von German
\caption[Georg Christoph Lichtenberg]{Georg Christoph Lichtenberg\footnotemark}
\end{ctsquote}
\footnotetext{%
  Georg Christoph Lichtenberg\index{Lichtenberg, Georg Christoph}, 
  German writer and physicist (1742-1799),\\ 
  (also see: \url{http://en.wikipedia.org/wiki/Georg\_Christoph\_Lichtenberg})
}

@@

A 0-1 knapsack problem can be solved by testing the $ 2^n $   possibilities for 
$ x_1, \dots, x_n $. The best method requires $ O(2^{n/2}) $  operations, which 
for $ n=100 $  with $ 2^{100} \approx 1.27 \cdot 10^{30} $  and  $ 2^{n/2} 
\approx 1.13 \cdot 10^{15} $ represents an insurmountable hurdle for computers.
However, for special $ a_1, \dots, a_n $   the solution is quite easy to find, 
e.g. for $ a_i = 2^{i-1}. $  The binary representation of $ g $ immediately 
delivers $ x_1, \dots, x_n$. In general, the a 0-1 knapsack problem can be 
solved easily if a \index{Permutation} permutation\footnote{%
A permutation\index{Permutation} $\pi$ of the numbers $1, \dots, n$ is a change in the order in which these numbers are listed. For example, a permutation $\pi$ of $(1,2,3)$ is $(3,1,2),$ i.e. $\pi(1) = 3$, $\pi(2) =1$ and $\pi(3) = 2$.} 
$ \pi $  of $ 1, \dots, n $  exists with $ a_{\pi (j)} > \sum_{i=1}^{j-1} 
a_{\pi(i)} $. If, in addition, $ \pi $ is the identity, i.e. $ \pi(i)=i $ for $ 
i=1,2,\dots,n, $ then the sequence $ a_1, \dots , a_n $ is said to be 
super-increasing.
Crypto procedure~\ref{knapsackalgo} solves the knapsack problem with a super-increasing 
sequence in the time of $ O(n). $
\begin{cryptoprocedure}
\begin{tabbing}
\hspace*{0.5cm} \= \hspace*{0.5cm} \= \hspace*{0.5cm} \= \kill
\>{\bf for} $ i=n $ {\bf to} 1 {\bf do} \\
\>\> {\bf if} $ T\geq a_i $ {\bf then}\\
\>\> \> $ T:=T-s_i $ \\
\>\>\> $ x_i:=1 $ \\
\>\> {\bf else} \\
\>\>\> $ x_i:=0 $ \\
\>{\bf if} $ T=0 $ {\bf then} \\
\>\> $ X:=(x_1, \dots, x_n) $ is the solution. \\
\>{\bf else} \\
\>\> No solution exists.
\end{tabbing}
\caption{Solving knapsack problems with super-increasing weights}
\label{knapsackalgo}
\end{cryptoprocedure}

@@

%--------------------------------------------------------------------
\hypertarget{RSAVerfahren}{}
\subsection[The RSA procedure]
{The RSA procedure\footnote{%
Please compare chapters \ref{rsabeweis}, ff.
}$^,$\footnote{%
Using CT1\index{CrypTool 1} you can gain practical experience
with the RSA procedure via the menu {\bf Indiv.Procedures \textbackslash{} 
RSA Cryptosystem \textbackslash{} RSA Demonstration}.\\
You can find RSA also in CT2\index{CrypTool 2} and JCT\index{JCrypTool}.
}
}
% LateX-Learnings: Mit \footnotemark statt \footnote
% stand der höhere Index unten zweimal (oben ok).
%
%  {The RSA procedure\footnotemark$^,$\footnotemark}
%  \footnotetext{
%  Please compare chapters \ref{rsabeweis}, ff.
%  }
%  \footnotetext{
%  Using CrypTool\index{CrypTool} you can gain practical experience
%  with the RSA procedure via the menu {\bf Indiv.Procedures \textbackslash{} 
%  RSA Cryptosystem \textbackslash{} RSA Demonstration}.
%  }
%  
%   in the article ``Introduction to Elementary Number Theory with Examples'',
%   chapter~\ref{patternsandstructures}
%   \hyperlink{RSABeweis}{Beweis des RSA-Verfahrens mit Euler-Fermat}
\index{RSA} \label{rsaverfahren}

@@

For the Euler phi function it holds that: 
\[
\phi (1)=1, \phi(2)=1,
\phi(3)=2, \phi (4)=2, \phi(6)=2, \phi (10)= 4, \phi (15)=8.
\]
For example, $ \phi (24)=8, $ because
\[
|\{ x <24 : \ggt (x,24) =1 \}| =|\{1,5,7,11,13,17,19,23\}|.
\]
If  $ p $ is a prime number, then $ \phi (p)= p-1. $ \\
If we know the various prime factors $ p_1, \dots , p_k $ of $ N, $ then 
\[
\phi(N) = N \cdot (1-\frac{1}{p_1}) \,
\cdots \, (1-\frac{1}{p_k}).\footnote{%
Further formulas for the Euler phi function are in
chapter~\ref{L-Euler-Function} ``\nameref{L-Euler-Function}''.
}
\]
Table~\ref{phi15} gives the values up to $15$. 
In the case of $ N=pq $, 
\[
\phi (N)= pq(1-1/p)(1-1/q) = p(1-1/p)q(1-1/q)=(p-
1)(q-1).
\]
\end{remark}

@@

\vskip +20 pt
%--------------------------------------------------------------------
\hypertarget{HT-Discrete-Logarithm-as-Basis}{}
\section[The discrete logarithm as basis for public key procedures]
           {The discrete logarithm as basis for public key procedures\footnotemark}
\footnotetext{%
    \index{NT, Learning Tool for Number Theory}%
    \index{Educational tool NT}%
    With the educational tool for number theory {\bf NT} you can play with the
    distribution of the discrete logarithm values and apply Shank's
    baby-step-giant-step method:
    See learning units 6.1-6.3, pages 1-6/6.\\
    NT can be called in CT1\index{CrypTool 1} via the menu path
    {\bf Indiv. Procedures \textbackslash{} Number Theory Interactive
    \textbackslash{} Learning Tool for Number Theory}.
    See appendix \ref{s:appendix-Learn-NT}.
}
\label{L-Discrete-Logarithm-as-Basis}
\index{Discrete logarithm}\index{Logarithm problem!discrete}
Discrete logarithms form the basis for a large
number of algorithms for public-key procedures.

@@

The current record\index{Logarithm problem!record} (as of April 2007) for calculating discrete logarithms was
established in February 2007 by the group Kleinjung, Franke and Bahr at
University of Bonn.\footnote{%
   \url{http://www.nabble.com/Discrete-logarithms-in-GF(p)-----160-digits-t3175622.html}}
Kleinjung calculated the discrete logarithm modulo a 160 digit prime number $p$
and generator $g$:
$$
\begin{array}{r@{\:}c@{\:}l}
p & = & \lfloor 10^{159}\pi\rfloor + 119849 \\
  & = & 314159265358979323846264338327950288419716939937510582097494 \\
  &   & 459230781640628620899862803482534211706798214808651328230664 \\
  &   & 7093844609550582231725359408128481237299\\
g & = & 2
\end{array}
$$
The discrete logarithms $k$ of the following integer $y$ was
determined:\footnote{The integer $y$ was chosen as the first 159 digits of the
Euler number $e$.}
$$
\begin{array}{r@{\:}c@{\:}l}
y & = & \lfloor 10^{159}e\rfloor \\
  & = & 271828182845904523536028747135266249775724709369995957496696 \\
  &   & 762772407663035354759457138217852516642742746639193200305992 \\
  &   & 1817413596629043572900334295260595630738 \\
k & = & \log_g(y)\mod p\\
  & = & 829897164650348970518646802640757844024961469323126472198531 \\
&   & 845186895984026448342666252850466126881437617381653942624307 \\
&   & 537679319636711561053526082423513665596
\end{array}
$$
The search was performed with GNFS method (General Number Field Sieve,
Index-Calculus) \index{General Number Field Sieve (GNFS)} and took about 17 CPU
years on 3.2 GHz Xeon machines.

%--------------------------------------------------------------------
\subsection[Diffie-Hellman key agreement]
{Diffie-Hellman key agreement\footnotemark}
\footnotetext{%
  With CT1\index{CrypTool 1} this exchange protocol has been
  visualized: you can execute the single steps with concrete numbers using 
  menu {\bf Indiv. Procedures \textbackslash{} Protocols \textbackslash{}
  Diffie-Hellman Demonstration}.\\
  In JCT\index{JCrypTool}\index{JCT} you can find it in the default perspective
  via the menu item {\bf Visuals \textbackslash{} Diffie-Hellman Key
  Exchange (EC)}.
}

### 2012-11-19 script/en/chapters/movies.tex
CT script: Minor changes (more consistent, more up-to-date)

\item[\textrm{[Isau1997]}] \index{Isau 1997}
    Ralf Isau, \\
    {\em The Museum of the stolen memories (original title: Das Museum
    der gestohlenen Erinnerungen)}, Thienemann-Verlag, 1997/2003. \\
    In this exciting novel the last part of the oracle can only be
    solved with the joined help of the computer community.\\
    The book got several awards and exists in 8 different languages,
    but not in English yet.\\

@@

CrypTool contains an interactive educational tool for elementary number theory,
called ``NT''.\footnote{%
    NT can be called in CrypTool\index{CrypTool} via the menu path
    {\bf Indiv. Procedures \textbackslash{} Number Theory Interactive
    \textbackslash{} Learning Tool for Number Theory}.
}

@@

This visualized educational tool was build with Authorware 4.

>>>>

\item[\textrm{[Isau2003]}] \index{Isau 1997}
    Ralf Isau, \\
    {\em The Museum of the stolen memories (original title: Das Museum
    der gestohlenen Erinnerungen)}, Thienemann-Verlag, 1997/2003. \\
    In this exciting novel the last part of the oracle can only be
    solved with the joined help of the computer community.\\
    The book got several awards and exists in 8 different languages,
    but not in English yet.\\

@@

CT1\index{CrypTool 1} contains an interactive educational tool for elementary
{\em number theory}, called ``NT''.\footnote{%
    NT can be called in CT1\index{CrypTool 1} via the menu path
    {\bf Indiv. Procedures \textbackslash{} Number Theory Interactive
    \textbackslash{} Learning Tool for Number Theory}.
}

@@

This visualized educational tool was build with Authorware 4.\footnote{%
    As Authorware is outdated and as the vendor didn't make tools available
    to easily port it to his successor products, the program ZT will not be
    further developed.
}

### 2012-11-19 script/en/chapters/numbertheory.tex
CT script: Minor changes (more consistent, more up-to-date)

\clearpage
% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%\vspace{4ex}
\begin{ctsquote}
    To let the possible happen, you again and again have to try the impossible.
\caption[Hermann Hesse]{Hermann Hesse\footnotemark}
\end{ctsquote}
\addtocounter{footnote}{0}\footnotetext{%
  Hermann Hesse, German/Swiss writer and Nobel Prize winner, July 2, 1877 $-$ August 9, 1962.\\}
% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

@@

%be_2005: Erzwingen, dass die Abb. noch in diesem Kapitel !

\noindent Below the last records are explained in more detail.
The two methods, GNFS and SNFS, used to do so are shortly illustrated at the 
following web pages:
\index{General Number Field Sieve (GNFS)}
\index{Special Number Field Sieve (SNFS)}
\vspace{-10pt}
\begin{itemize}
\item[]
      \url{http://en.wikipedia.org/wiki/Special_number_field_sieve} \\
      \url{http://en.wikipedia.org/wiki/General_number_field_sieve}
\end{itemize}
%\vspace{-10pt}

% --------------------------------------------------------------------------
\vskip +20pt
\paragraph*{RSA-155} \label{RSA-155} \index{RSA-155}\mbox{}

@@

>>>>

\clearpage
% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%\vspace{4ex}
\begin{ctsquote}
    To let the possible happen, you again and again have to try the impossible.
\caption[Hermann Hesse]{Hermann Hesse\footnotemark}
\end{ctsquote}
\addtocounter{footnote}{0}\footnotetext{%
  Hermann Hesse, German/Swiss writer and Nobel Prize winner, July 2, 1877 $-$ August 9, 1962.}
% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

@@

Experiences about the ellipsed time of factorization\index{factorization}
with the open source software Pari-GP, Sage, CrypTool 1 and CrypTool 2)
can be found in ``Zeitexperimente zur Faktorisierung''\footnote{%
R.-H.Schulz und H.Witten: ``Zeitexperimente zur Faktorisierung''. Ein Beitrag
zur Didaktik der Kryptographie'', in LogIn Heft Nr. 166/167 (2010) 113-120,
(available currently only in German)\\
\url{http://bscw.schule.de/pub/bscw.cgi/d864899/Schulz_Witten_Zeit-Experimente.pdf}
}.

\vskip +20pt
%\noindent
Below these last records are explained in more detail\footnote{%
The two methods, GNFS and SNFS, used to do so are shortly illustrated at the 
following web pages:
\index{General Number Field Sieve (GNFS)}
\index{Special Number Field Sieve (SNFS)}
\begin{itemize}
\item[]
      \url{http://en.wikipedia.org/wiki/Special_number_field_sieve} \\
      \url{http://en.wikipedia.org/wiki/General_number_field_sieve}
\end{itemize}
\vspace{-10pt}
}:

% --------------------------------------------------------------------------
%\vskip +20pt
\paragraph*{RSA-155} \label{RSA-155} \index{RSA-155}\mbox{}

@@

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\newpage
\hypertarget{NumberTheory_Appendix_D2_Koblenz}{}
\section{Appendix: Interactive presentation about the RSA cipher}
\label{l:NumberTheory_Appendix_D2_Koblenz}{}

The folowing presentation (last update Nov. 2010) shows the basics of the
RSA cipher in an interactive way.\\
There are three variants:
\begin{itemize}
 \item Powerpoint 2007 (for download; dynamical, animated)\footnote{%
    \url{http://www.cryptool.org/images/ct1/presentations/RSA/RSA-de.pptx}}
 \item PDF (for download; static, no interaction)\footnote{%
    \url{http://www.cryptool.org/images/ct1/presentations/RSA/RSA-de(keine%20Interaktivitaet).pdf}}
 \item Flash (can be started within the browser, requires JavaScript; time-controlled replay)\footnote{%
\url{http://www.cryptool.org/images/ct1/presentations/RSA/RSA-Flash-de/player.html}}
\end{itemize}

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.4]{figures/Interactive_RSA_Presentation_E.png}
\caption{Screenshot RSA Presentation} 
\label{l_Interactive_RSA_Presentation}
\end{center}
\end{figure}

### 2012-11-19 script/en/chapters/what-is-sage.tex
CT script: Minor changes (more consistent, more up-to-date)

\noindent You can get a quick impression of Sage e.g. with David Joyner's
``Invitation to Sage''\footnote{%
\url{http://sage.math.washington.edu/home/wdj/teaching/calc1-sage/an-invitation-to-sage.pdf}\\
(last update 2009).}.

\noindent The official Sage online documentation\footnote{%
  The corresponding official PDF documents can be downloaded at\\
  \url{http://www.sagemath.org/help.html}
} is available at: \url{http://www.sagemath.org/doc}.

@@

\noindent Especially David Kohel's notes at\\
{\centering \url{http://www.sagemath.org/library/crypto.pdf} }
or the same eventually newer at\\
{\centering \url{http://sage.math.washington.edu/home/wdj/teaching/kohel-crypto.pdf} }\\
can be used to teach such a course that incorporates Sage.

% ---------------------------------------------------------------------------
\section*{Sage user interfaces}
Sage is available free of charge and can be downloaded from the
following website:
\begin{center}
  \url{http://www.sagemath.org} \\
\end{center}
The default interface to Sage is command line based, as shown in
figure~\ref{fig:sage_cmd_interfaces}. However, there is a
graphical user interface to the software as well in the form of the
Sage notebook (see figure~\ref{fig:sage_gui_interfaces}). We can even
use Sage notebooks\footnote{%
Further details about Sage notebooks can be found at 
chapter~\ref{ec:Sage_Massierer} 
(``\nameref{ec:Implementing-for-Education}''
$\Rightarrow$ ``\nameref{ec:Sage_Massierer}'')
                      }
online at different servers, without having to install Sage locally, e.g:
\begin{center}
\url{http://www.sagenb.org} or\\
\url{http://sage.mathematik.uni-siegen.de:8000}
\end{center}
.

Sage runs under many Linux distributions, Mac OS X, and Windows.
For the Windows platform, a complete distribution of Sage currently
only runs as a VMware image. However, a full native port of Sage
to Windows is currently in progress.

>>>>

\noindent You can get a quick impression of Sage e.g. with the references in this footnote\footnote{%
- ``Invitation to Sage'' by David Joyner, last update 2009\\
\url{http://sage.math.washington.edu/home/wdj/teaching/calc1-sage/an-invitation-to-sage.pdf}\\
- ``The SDSU Sage Tutorial'',\\
  \url{http://www-rohan.sdsu.edu/~mosulliv/sagetutorial/}\\
  \url{http://www-rohan.sdsu.edu/~mosulliv/sagetutorial/sagecalc.html}\\
- ``SAGE For Newbies'' by Ted Kosan, 2007,\\
   \url{http://sage.math.washington.edu/home/tkosan/newbies_book/sage_for_newbies_v1.23.pdf}
}.

\noindent The official Sage online documentation\footnote{%
  The corresponding official PDF documents can be downloaded at\\
  \url{http://www.sagemath.org/help.html}, \url{http://www.sagemath.org/doc} and \url{http://planet.sagemath.org}.
} is available at: \url{http://www.sagemath.org}.

@@

\noindent Comprehensive introductions into cryptography are in this
footnote\footnote{%
- David Kohel's notes from 2008 are a ready cource\\
{\centering \url{http://www.sagemath.org/library/crypto.pdf} }
or the same eventually newer at\\
{\centering \url{http://sage.math.washington.edu/home/wdj/teaching/kohel-crypto.pdf} }.\\
- ``Introduction to Cryptography with Open-Source Software, a very good book
  from Alasdair McAndrew, CRC, 2011\\
}.

% ---------------------------------------------------------------------------
\section*{Sage user interfaces}
Sage is available free of charge and can be downloaded from the
following website:
\begin{center}
  \url{http://www.sagemath.org} \\
\end{center}
The default interface to Sage is {\bf command line} based, as shown in
figure~\ref{fig:sage_cmd_interfaces}. However, there is a
graphical user interface to the software as well in the form of the
Sage notebook (see figure~\ref{fig:sage_gui_interfaces}). We can even
use Sage {\bf notebooks}\footnote{%
Further details about Sage notebooks can be found at 
chapter~\ref{ec:Sage_Massierer} 
(``\nameref{ec:Implementing-for-Education}''
$\Rightarrow$ ``\nameref{ec:Sage_Massierer}'').
                      }
online at different servers, without having to install Sage locally, e.g:
\begin{center}
\url{http://www.sagenb.org} or\\
\url{http://sage.mathematik.uni-siegen.de:8000}
\end{center}

Sage runs under many Linux distributions, Mac OS X, and Windows.
For the Windows platform, a complete distribution of Sage currently
only runs as a VMware image.
